VFTS682: a confirmed dynamical ejection?

### 2018-05-09 VFTS682.tex
minor

VFTS682 is labeled in the Gaia DR2
catalog\footnote{\url{https://gea.esac.esa.int/archive/}} with the
source id 4657685637907503744. The star has a
\texttt{visibility\_period} = 17, which counts how many observations have
been used to reconstruct its astrometric solution \citep[][]{lindengren:18}. Its G-band
magnitude is 15.65, cf. the V-band magnitude of 16.08
\citep[][]{evans:11, bestenlehner:11}, and the reported
\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
data for VFTS682 are trustworthy. However, the effective temperature
reported is one order of magnitude lower than what found by
\cite{bestenlehner:11}, and the best fit parallax of this star is
negative. We do not use the effective temperature of the star anywhere
in this study, and we attribute the unphysical value of the parallax
to the large distance to the LMC. Our main findings do not rely on the
parallax value reported in the Gaia catalog.

@@

To compare the astrometry of VFTS682 and derive its peculiar motion,
we then select data from the Gaia DR2 catalog for two regions: the
``surroundings'' of VFTS 682, and the ``R136 cluster''. The
surrounding region is defined by all the stars in a target of 10 arcminutes around
VFTS682 fulfilling the following criteria: we require a G-band
magnitude brighter than 17, to mimic the completeness threshold for
the VFTS survey, \texttt{visibility\_period} >= 5, and
\texttt{astrometric\_excess\_noise} <= 1. This selection yields
roughly 17800 stars, including the subset of stars which we consider to be
part of R136, discussed below. We further require  these stars to have finite proper
motion components and relative errors, reducing further the sample to
14856 stars. % A more strict sample selection for the surroundings can
% be found in \cite{lennon:18}.

The ``R136 cluster'' is effectively defined by taking all the stars
within 25 arcseconds from R136a, one of the most massive members of
the cluster itself \citep[][]{crowther:10}, requiring the same
``quality'' criteria applied above. Because of the magnitude cut
however, this would yield an incredibly small sample of 15 stars only
% which makes no sense at all
\todo{improve definition
  of stars from R136, describe accordingly -- check with Danny}

Throughout this study, we assume the same distance of $50$\,kpc to the star, and to
the 30 Doradus region as a whole, since the parallax for VFTS682
listed in the Gaia DR2 catalog is negative.

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/main_plot}  
  \caption{position and projected relative velocity to
    R136. \todo{load pretty picture on background, check scale, clean
      foreground stars with parallax, add cone of uncertainty}}
  \label{fig:main}
\end{figure*}

@@

Random notes: vsini<200 form \cite{schneider:18}, age $1\pm 0.2$\,Myr
from \cite{schneider:18}

>>>>

VFTS682 is labeled in the Gaia DR2
catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}} with the
source id 4657685637907503744. The star has a
\texttt{visibility\_period} = 17, which counts how many observations have
been used to reconstruct its astrometric solution
\citep[][]{lindengren:18}. Its reported G-band
magnitude is 15.65, cf. the V-band magnitude of 16.08
\citep[][]{evans:11, bestenlehner:11}, and the reported
\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
data for VFTS682 are trustworthy. However, the effective temperature
reported in Gaia DR2 is one order of magnitude lower than what found by
\cite{bestenlehner:11}, and the best fit parallax of this star is
negative. We do not use the effective temperature of the star anywhere
in this study, and we attribute the unphysical value of the parallax
to the large distance to the LMC. Our main findings do not rely on the
parallax nor the effective temperature values reported in the Gaia DR2
catalog.

%% should I really mention the line of sight velocity? Not really
%% using it...

@@

To compare the astrometry of VFTS682 and derive its peculiar motion,
we then select data from the Gaia DR2 catalog for two regions: the
``surroundings'' of VFTS 682, and the ``R136 cluster''. The
surrounding region is defined by all the stars in a target of 10 arcminutes around
VFTS682 fulfilling the following criteria: we require a G-band
magnitude brighter than 17, to mimic the completeness threshold for
the VFTS survey, \texttt{visibility\_period} >= 5, and
\texttt{astrometric\_excess\_noise} <= 1. This selection yields
roughly 1330 stars, including the subset of stars which we consider to be
part of R136, discussed below. We further require  these stars to have finite proper
motion components and relative errors, reducing further the sample to
1323 stars. % A more strict sample selection for the surroundings can
% be found in \cite{lennon:18}.

The ``R136 cluster'' is effectively defined by taking all the stars
within 25 arcseconds from R136a, one of the most massive members of
the cluster itself \citep[][]{crowther:10}, requiring the same
``quality'' criteria applied above. This selection yields 62 stars.
\todo{improve definition
  of stars from R136, describe accordingly -- check with Danny}

Throughout this study, we assume the same distance of $50$\,kpc to the star, and to
the 30 Doradus region as a whole, since the parallax for VFTS682
listed in the Gaia DR2 catalog is negative. %% this is what makes the
                                %% LOS velocity useless...how to improve?

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/main_plot}  
  \caption{position and projected relative velocity to
    R136. \todo{orient properly, load  picture on background, add cone of uncertainty}}
  \label{fig:main}
\end{figure*}

@@

Random notes: $v\sin(i)<200\,\mathrm{km\ s^{-1}}$ form \cite{schneider:18}, age $1.0\pm 0.2$\,Myr
from \cite{schneider:18}
